---
title: "Dropout event recovery"
output: html_notebook
---
### Setup knitr and load utility functions
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir="E:/DISC/reproducibility")
```
```{r}
utilities_path = "./source/utilities.r"
source(utilities_path)
```
### Load raw data and downsampling
Here, we use MELANOMA dataset as an example.</br>
The fraction of sampling is set to 50% of the original library size across cells.</br>
Only 1 replicate will be generated here as asn example.</br>
Only imputed genes will be kept for comparison.
```{r}
raw_data = readh5_loom("./data/melanoma/raw.loom")
ds_dir = "./data/melanoma/ds_0.5"
#dir.create(ds_dir, showWarnings = F, recursive = T)
#observed_data = downsampling_cell(0.5, raw_data)
#save_h5(paste0(ds_dir, "/observed.loom"), t(observed_data))
observed_data = readh5_loom(paste0(ds_dir, "/observed.loom"))
expressed_cell = rowSums(observed_data > 0)
gene_expression = rowSums(observed_data)
gene_filter = expressed_cell >= 10 & gene_expression > expressed_cell * 1
raw_data = raw_data[gene_filter, ]
observed_data = observed_data[gene_filter, ]
used_genes = rownames(observed_data)
print(dim(raw_data))
print(dim(observed_data))
data_list = list(Raw = raw_data, Observed = observed_data)
rm(raw_data, observed_data)
```
### Load downsampling data and imputation results
We <a href="https://github.com/iyhaoo/DISC/blob/master/reproducibility/tutorials/run_imputation.md">run imputation</a>  using downsampling data.</br>
Here, we load their results.
```{r}
data_list[["DISC"]] = readh5_imputation(paste0(ds_dir, "/DISC.hdf5"), with_outliers = T)[used_genes, ]
print(dim(data_list[["DISC"]]))
data_list[["SAVER"]] = readh5_imputation(paste0(ds_dir, "/SAVER.hdf5"))
data_list[["MAGIC"]] = readh5_imputation(paste0(ds_dir, "/MAGIC.hdf5"))
data_list[["DCA"]] = readh5_imputation(paste0(ds_dir, "/DCA.hdf5"))
data_list[["scScope"]] = readh5_imputation(paste0(ds_dir, "/scScope.hdf5"))
data_list[["scVI"]] = readh5_imputation(paste0(ds_dir, "/scVI.hdf5"))
cell_number = ncol(data_list[["Raw"]])
gene_number = length(used_genes)
```
### Settings
```{r}
method_names = setdiff(names(data_list), "Raw")
method_color = c("gray80", "red", "blue4", "yellow4", "green", "purple", "cyan")
names(method_color) = method_names
text_color = rep("black", length(method_names))
names(text_color) = method_names
text_color["DISC"] = "red"
bar_color = rep("gray50", length(method_names))
names(bar_color) = method_names
bar_color["Raw"] = "gray80"
bar_color["DISC"] = "red"
```
### MAE
```{r}
#output_dir = paste0(ds_dir, "/MAE")
#dir.create(output_dir, showWarnings = F, recursive = T)
mae_gt0 = c()
mae_eq0 = c()
scale_factor = 1 / 0.5
for(this_method in method_names){
  eq0 = sum(data_list[["Raw"]] > 0 & data_list[["Observed"]] == 0)
  gt0 = sum(data_list[["Raw"]] > 0 & data_list[["Observed"]] > 0)
  sae_gt0 = sum(sapply(used_genes, function(x){
    expressed_mask = data_list[["Raw"]][x, ] > 0 & data_list[["Observed"]][x, ] > 0
    return(sum(abs(data_list[["Raw"]][x, expressed_mask] - (data_list[[this_method]][x, expressed_mask] * scale_factor))))
  }))
  sae_eq0 = sum(sapply(used_genes, function(x){
    expressed_mask = data_list[["Raw"]][x, ] > 0 & data_list[["Observed"]][x, ] == 0
    return(sum(abs(data_list[["Raw"]][x, expressed_mask] - (data_list[[this_method]][x, expressed_mask] * scale_factor))))
  }))
  mae_gt0 = c(mae_gt0, sae_gt0 / gt0)
  mae_eq0 = c(mae_eq0, sae_eq0 / eq0)
  print(this_method)
}
names(mae_gt0) = method_names
names(mae_eq0) = method_names
```

```{r fig.height=5, fig.width=4.5}
barplot_usage(mae_eq0, main = "Zero entries", cex.main = 1.5, bar_color = bar_color, text_color = text_color, use_data_order = T, ylab = "Log (MAE + 1)", cex.lab = 1.5, font.main = 1)
```

### CMD
```{r}
output_dir = paste0(ds_dir, "/CMD")
dir.create(output_dir, showWarnings = F, recursive = T)
vst_file = paste0(output_dir, "/vst_gene.tsv")
if(file.exists(vst_file)){
  hvg_info = read.table(vst_file)
  print("load vst_file")
}else{
  hvg_info = FindVariableFeatures_vst_by_genes(data_list[["Raw"]])
  hvg_info = hvg_info[order(hvg_info$variance.standardized, decreasing = T), ]
  write.table(hvg_info, paste0(output_dir, "/vst_gene.tsv"), sep = "\t", quote = F, row.names = T, col.names = T)
}
used_feature_genes = rownames(hvg_info)[1:300]
cor_all = list()
for(method in names(data_list)){
  cor_all[[method]] = matrix(nrow = length(used_feature_genes), ncol = length(used_feature_genes), dimnames = list(used_feature_genes, used_feature_genes))
  this_mat = delete_lt0.5(data_list[[method]])[used_feature_genes, ]
  no_cores <- detectCores() - 1
  cl <- makeCluster(no_cores)
  clusterExport(cl, varlist = c("this_mat", "method"))
  return_list = parLapply(cl, seq(nrow(this_mat) - 1), function(x){
    return_vector = rep(NA, nrow(this_mat) - x)
    ii_express = this_mat[x, ]
    ii_mask = ii_express > 0
    for(jj in (x + 1):nrow(this_mat)){
      jj_express = this_mat[jj, ]
      if(method %in% c("")){
        express_mask = rep(T, length(ii_mask))
      }else{
        express_mask = ii_mask | jj_express > 0
      }
      if(sum(express_mask) > 0){
        this_corr = cor(ii_express[express_mask], jj_express[express_mask], use = "pairwise.complete.obs")
        if(is.na(this_corr)){
          return_vector[jj - x] = 0
        }else{
          return_vector[jj - x] = this_corr
        }
      }else{
        return_vector[jj - x] = 0
      }
    }
    return(list("return_vector" = return_vector))
  })
  stopCluster(cl)
  cor_all[[method]][1, 1] = 1
  for(jj in 1:length(return_list)){
    cor_all[[method]][jj, (jj + 1): nrow(this_mat)] = return_list[[jj]][["return_vector"]]
    cor_all[[method]][(jj + 1): nrow(this_mat), jj] = return_list[[jj]][["return_vector"]]
    cor_all[[method]][(jj + 1), (jj + 1)] = 1
  }
  print(method)
}
saveRDS(cor_all, paste0(output_dir, "/cor_all.rds"))
cmd_result = c()
for(method in method_names){
  cmd_result = c(cmd_result, calc_cmd(cor_all[["Raw"]], cor_all[[method]]))
}
names(cmd_result) = method_names
print(cmd_result)
```
```{r fig.height=5, fig.width=4.5}
barplot_usage(cmd_result, main = "", cex.main = 1.5, bar_color = bar_color, text_color = text_color, use_data_order = T, ylab = "CMD", cex.lab = 1.5, font.main = 1, ylim = c(-0.1, 1))
```


### Gene correlation
```{r}
gene_correlation_mat = matrix(nrow = gene_number, ncol = length(method_names), dimnames = list(used_genes, method_names))

for(method in method_names){
  for(ii in used_genes){
  	raw_expression = data_list[["Raw"]][ii, ]
  	raw_expressed_mask = raw_expression != 0
  	raw_expressed_entries = raw_expression[raw_expressed_mask]
  	if(length(table(raw_expressed_entries)) > 1 & length(raw_expressed_entries) >= (0.1 * cell_number)){
    	gene_correlation_mat[ii, method] <- cor(raw_expressed_entries, data_list[[method]][ii, raw_expressed_mask], method = "pearson")
  	}
  }
  print(method)
}
print(colMeans(gene_correlation_mat, na.rm = T))
```
```{r fig.height=5, fig.width=4.5}
barplot_usage(colMeans(gene_correlation_mat, na.rm = T), main = "", cex.main = 1.5, bar_color = bar_color, text_color = text_color, use_data_order = T, decreasing = T, ylab = "Gene correlation with reference", cex.lab = 1.5, font.main = 1, ylim = c(-0.1, 1))
```
### Cell correlation
```{r}
cell_correlation_mat = matrix(nrow = cell_number, ncol = length(method_names), dimnames = list(c(), method_names))
for(method in method_names){
  for(ii in seq(cell_number)){
  	raw_expression = data_list[["Raw"]][, ii]
  	raw_expressed_mask = raw_expression != 0
  	raw_expressed_entries = raw_expression[raw_expressed_mask]
  	if(length(table(raw_expressed_entries)) > 1 & length(raw_expressed_entries) >= (0.1 * gene_number)){
    	cell_correlation_mat[ii, method] <- cor(raw_expressed_entries, data_list[[method]][raw_expressed_mask, ii], method = "pearson")
  	}
  }
  print(method)
}
print(colMeans(cell_correlation_mat, na.rm = T))
```
```{r fig.height=5, fig.width=4.5}
barplot_usage(colMeans(cell_correlation_mat, na.rm = T), main = "", cex.main = 1.5, bar_color = bar_color, text_color = text_color, use_data_order = T, decreasing = T, ylab = "Cell correlation with reference", cex.lab = 1.5, font.main = 1, ylim = c(-0.1, 1))
```







